CF Registry Analysis

### 2023-09-29 Christine Chan/CF Registry/cf_registry_analysis.qmd
cf registry analyses updated

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
editor: source
---

@@

annual <- read.csv("~/Documents/Work/Christine Chan/CF registry data/CFF19_Annualized_Del1.csv",na.strings = c("", "N/A")

@@

annual = annual %>% filter(cfrd_status_annual != 4,cfrd_status_annual != 5)
annual$cfrd_status_annual = factor(annual$cfrd_status_annual,levels = 1:3,
                                   labels = c("CFRD-","CFRD-","CFRD+"))
demogs = read.csv("~/Documents/Work/Christine Chan/CF registry data/CFF19_DemogCFDiag_Del1.csv")
demogs = demogs %>% select(eDWID,Gender,MutClass)
annual = left_join(annual,demogs,by = join_by(eDWID))

@@

log_reg = function(year){

@@

  df = annual %>% 
    filter(ReviewYear==year) %>%
    select(eDWID,Age_YrEnd,Gender,MutClass,A_IsOnEnzymes,A_FEV1,A_FEV1_pct_predicted,
           A_bmivalue,A_bmipercentile,A_height,A_weight,A_heightpercentile,
           A_supplement_feeding,A_feeding2,
           NumPulmExacerbation,A_pseudomonasaeruginosa,
           A_fungalyeast1,A_pulmonarycomplications1,A_mycocultureresults1,
           A_corticosteroids1,cfrd_status_annual,A_Vx770,A_VX809comb)

@@

  df$A_feeding2[df$A_supplement_feeding == 0] = 0

@@

  df$bmi = df$A_weight / ((df$A_height/100)^2)
  df$age_perc = ifelse(df$Age_YrEnd < 20,df$Age_YrEnd,20)
  df$height_perc = sds(df$A_height,age = df$age_perc,
                    sex = df$Gender,male = "M",female = "F",
                    item = "height2_20",ref = cdc.ref,type = "perc")*100
  df$bmi_perc = sds(df$bmi,age = df$age_perc,
                    sex = df$Gender,male = "M",female = "F",
                    item = "bmi",ref = cdc.ref,type = "perc")*100

@@

  df$MutClass = factor(df$MutClass,levels = c("1-3","4-5","Oth","N/A"),
                       labels = c("1-3","4-5","Oth","Unknown"))
  df$NumPulmExacerbation=cut(df$NumPulmExacerbation,labels = c("0","1","2",">=3"),
                             breaks = c(0,1,2,3,Inf),right = F)
  df$A_FEV1_pct_predicted = cut(df$A_FEV1_pct_predicted,
                                breaks = c(0,40,80,100,Inf),right = F)
  df$A_FEV1_pct_predicted = relevel(df$A_FEV1_pct_predicted,ref = "[100,Inf)")
  df$height_perc = cut(df$height_perc,breaks = c(0,10,25,50,Inf))
  df$height_perc = relevel(df$height_perc,ref = "(50,Inf]")
  df$bmi_perc = cut(df$bmi_perc,breaks = c(0,10,50,85,Inf))
  df$bmi_perc = relevel(df$bmi_perc,ref = "(50,85]")
  df$age_group = cut(df$Age_YrEnd,breaks = c(0,10,18,30,Inf))
  df$age_group = relevel(df$age_group,ref = "(10,18]")

@@

  bugs = c("A_pseudomonasaeruginosa","A_fungalyeast1",
           "A_pulmonarycomplications1","A_mycocultureresults1")

@@

  yn_vars = c("A_IsOnEnzymes","A_corticosteroids1","A_feeding2",
              "A_Vx770","A_VX809comb",bugs)
  df[,yn_vars] = lapply(df[,yn_vars],factor,levels = 0:1,labels = c("No","Yes"))

@@

  empty = which(colSums(is.na(df))==nrow(df))
  if(length(empty)>0){
    df = df[,-empty]

@@

  vars = colnames(df)[which(!colnames(df) %in% 
                              c(bugs,"cfrd_status_annual","eDWID","A_bmivalue",
                                "A_bmipercentile","A_height","A_heightpercentile",
                                "Age_YrEnd","A_weight","bmi", "age_perc",
                                "A_supplement_feeding"))]
  base_mod = as.formula(paste0("cfrd_status_annual~",paste0(vars,collapse = "+")))

@@

  mod1 = glm(base_mod,data = df,family = "binomial")

@@

  print(forest_model(mod1))

@@

  if ("A_pseudomonasaeruginosa" %in% colnames(df)){
    mod2 = glm(update(base_mod,.~.+A_pseudomonasaeruginosa),data = df,family = "binomial")

@@

    print(forest_model(mod2))

@@

  if("A_fungalyeast1" %in% colnames(df)){
    mod3 = glm(update(base_mod,.~.+A_fungalyeast1),data = df,family = "binomial")

@@

    print(forest_model(mod3))

@@

  if("A_pulmonarycomplications1" %in% colnames(df)){
    mod4 = glm(update(base_mod,.~.+A_pulmonarycomplications1),data = df,family = "binomial")

@@

    print(forest_model(mod4))

@@

  if("A_mycocultureresults1" %in% colnames(df)){
    mod5 = glm(update(base_mod,.~.+A_mycocultureresults1),data = df,family = "binomial")

@@

    print(forest_model(mod5))

>>>>

---
title: "CF Registry Analysis"
author: "Tim Vigers"
date: "today"
date-format: long
format:
  html:
    toc: true
    toc-depth: 5
    toc-float: true
    code-fold: true
    self-contained: true
    fig-cap-location: top
    page-layout: full
execute:
  echo: false
  warning: false
editor: source
---

@@

annual <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_Annualized_Del1.csv",
  na.strings = c("", "N/A")

@@

annual <- annual %>% filter(cfrd_status_annual != 4, cfrd_status_annual != 5)
annual$cfrd_status_annual <- factor(annual$cfrd_status_annual,
  levels = 1:3,
  labels = c("CFRD-", "CFRD-", "CFRD+")
demogs <- read.csv("/Users/timvigers/Dropbox/Work/Christine Chan/CF registry data/Data_Raw/CFF19_DemogCFDiag_Del1.csv")
demogs <- demogs %>% select(eDWID, Gender, MutClass)
annual <- left_join(annual, demogs, by = join_by(eDWID))

@@

log_reg <- function(year) {

@@

  df <- annual %>%
    filter(ReviewYear == year, Age_YrEnd >= 10) %>%
    select(
      eDWID, Age_YrEnd, Gender, MutClass, A_IsOnEnzymes, A_FEV1, A_FEV1_pct_predicted,
      A_bmivalue, A_bmipercentile, A_height, A_weight, A_heightpercentile,
      A_supplement_feeding, A_feeding2:A_feeding4,
      NumPulmExacerbation, A_pseudomonasaeruginosa,
      A_fungalyeast1, A_pulmonarycomplications1, A_mycocultureresults3,
      A_corticosteroids1, cfrd_status_annual, A_Vx770, A_VX809comb,
      A_hepatobiliary1_3, contains("A_hepatobiliary2_")

@@

  # Include all kinds of tube feeds
  df$A_feeding2[df$A_supplement_feeding == 0] <- 0
  df$A_feeding3[df$A_supplement_feeding == 0] <- 0
  df$A_feeding4[df$A_supplement_feeding == 0] <- 0
  df$tube_feeding <-
    apply(df[, c("A_feeding2", "A_feeding3", "A_feeding4")], 1, function(r) {
      if (all(is.na(r))) {
        return(NA)
      } else {
        return(max(as.numeric(r), na.rm = T))
    })
  # Liver disease
  liver <- c("A_hepatobiliary1_3", paste0("A_hepatobiliary2_", 1:4))
  df$liver_disease <-
    apply(df[, liver], 1, function(r) {
      if (all(is.na(r))) {
        return(NA)
      } else {
        return(max(as.numeric(r), na.rm = T))
    })

@@

  df$bmi <- df$A_weight / ((df$A_height / 100)^2)
  df$age_perc <- ifelse(df$Age_YrEnd < 20, df$Age_YrEnd, 20)
  df$height_perc <- sds(df$A_height,
    age = df$age_perc,
    sex = df$Gender, male = "M", female = "F",
    item = "height2_20", ref = cdc.ref, type = "perc"
  ) * 100
  df$bmi_perc <- sds(df$bmi,
    age = df$age_perc,
    sex = df$Gender, male = "M", female = "F",
    item = "bmi", ref = cdc.ref, type = "perc"
  ) * 100

@@

  df$MutClass <- factor(df$MutClass,
    levels = c("1-3", "4-5", "Oth", "N/A"),
    labels = c("1-3", "4-5", "Oth", "Unknown")
  df$NumPulmExacerbation <- cut(df$NumPulmExacerbation,
    labels = c("0", "1", "2", ">=3"),
    breaks = c(0, 1, 2, 3, Inf), right = F
  df$A_FEV1_pct_predicted <- cut(df$A_FEV1_pct_predicted,
    breaks = c(0, 40, 80, 100, Inf), right = F
  df$A_FEV1_pct_predicted <- relevel(df$A_FEV1_pct_predicted, ref = "[100,Inf)")
  df$height_perc <- cut(df$height_perc, breaks = c(0, 10, 25, 50, Inf))
  df$height_perc <- relevel(df$height_perc, ref = "(50,Inf]")
  df$bmi_perc <- cut(df$bmi_perc, breaks = c(0, 10, 50, 85, Inf))
  df$bmi_perc <- relevel(df$bmi_perc, ref = "(50,85]")
  df$age_group <- cut(df$Age_YrEnd, breaks = c(10, 18, 30, Inf), right = F)
  df$age_group <- relevel(df$age_group, ref = "[10,18)")

@@

  bugs <- c(
    "A_pseudomonasaeruginosa", "A_fungalyeast1",
    "A_pulmonarycomplications1", "A_mycocultureresults3"

@@

  yn_vars <- c(
    "A_IsOnEnzymes", "A_corticosteroids1", "A_feeding2",
    "A_Vx770", "A_VX809comb", bugs, "tube_feeding", "liver_disease"
  df[, yn_vars] <- lapply(df[, yn_vars], factor, levels = 0:1, labels = c("No", "Yes"))
  df$A_IsOnEnzymes <- relevel(df$A_IsOnEnzymes, ref = "Yes")

@@

  empty <- which(colSums(is.na(df)) == nrow(df))
  if (length(empty) > 0) {
    df <- df[, -empty]

@@

  vars <- colnames(df)[which(!colnames(df) %in%
    c(
      bugs, "cfrd_status_annual", "eDWID", "A_bmivalue",
      "A_bmipercentile", "A_height", "A_heightpercentile",
      "Age_YrEnd", "A_weight", "bmi", "age_perc",
      "A_supplement_feeding", "A_feeding2", "A_feeding3",
      "A_feeding4", "A_FEV1", liver
    ))]
  base_mod <- as.formula(paste0("cfrd_status_annual~", paste0(vars, collapse = "+")))
  # Labels
  labels <- list(
    cfrd_status_annual = "CFRD Status", MutClass = "Mutation class",
    Gender = "Sex", A_IsOnEnzymes = "Pancreatic enzyme use",
    A_FEV1_pct_predicted = "FEV1 percent predicted",
    tube_feeding = "Enteric feedings", liver_disease = "Liver disease",
    A_corticosteroids1 = "Use of corticosteroids",
    NumPulmExacerbation = "Admissions for PEx", age_group = "Age group",
    height_perc = "Height percentile", bmi_perc = "BMI percentile",
    A_Vx770 = "Use of CFTR modulator Vx770",
    A_VX809comb = "Use of CFTR modulator Vx809 combination",
    A_pseudomonasaeruginosa = "Pseudomonas aeruginosa",
    A_fungalyeast1 = "Aspergillus (any species)",
    A_pulmonarycomplications1 = "ABPA",
    A_mycocultureresults3 = "Microorganisms in myco culture"
  var_label(df[, vars[which(vars %in% colnames(df))]]) <-
    as.character(labels[vars[which(vars %in% colnames(df))]])
  var_label(df[, bugs[which(bugs %in% colnames(df))]]) <-
    as.character(labels[bugs[which(bugs %in% colnames(df))]])

@@

  mod1 <- glm(base_mod, data = df, family = "binomial")

@@

  print(forest_model(mod1, exclude_infinite_cis = F))

@@

  if ("A_pseudomonasaeruginosa" %in% colnames(df)) {
    mod2 <- glm(update(base_mod, . ~ . + A_pseudomonasaeruginosa), data = df, family = "binomial")

@@

    print(forest_model(mod2, exclude_infinite_cis = F))

@@

  if ("A_fungalyeast1" %in% colnames(df)) {
    mod3 <- glm(update(base_mod, . ~ . + A_fungalyeast1), data = df, family = "binomial")

@@

    print(forest_model(mod3, exclude_infinite_cis = F))

@@

  if ("A_pulmonarycomplications1" %in% colnames(df)) {
    mod4 <- glm(update(base_mod, . ~ . + A_pulmonarycomplications1), data = df, family = "binomial")

@@

    print(forest_model(mod4, exclude_infinite_cis = F))

@@

  if ("A_mycocultureresults3" %in% colnames(df)) {
    mod5 <- glm(update(base_mod, . ~ . + A_mycocultureresults3), data = df, family = "binomial")

@@

    print(forest_model(mod5, exclude_infinite_cis = F))
